Title here

### 2017-01-04 portfolio/text/ms.Rmd
Edits

Individuals who rely on the extraction of natural resources for their
livelihood --- fishers, farmers, miners, and loggers, among others --- face
high levels of income variability driven by a mix of environmental, biological,
and economic variability. A key component to managing these industries is
identifying how regulatory actions and individual behaviour affect income
variability and by extension the economic stability and sustainable use of
natural resources. For commercial fisheries, recent work has identified that
communities and vessels that fish a greater diversity of species have less
revenue variability than those that fish a lower diversity. However, it
is unclear if these benefits extend to the level of the individual and how
year-to-year changes in diversification affect revenue and revenue variability.
Here we identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can diversify fishing activities.
We show that fishing permits with higher species diversity reduces individual
revenue variability and fishing an additional permit is associated with higher
revenue and lower variability. However, increasing species diversity within the
constraints of a permit set has a fishery-dependent effect on revenue and is
usually (`r g2_dens`% probability) associated with increased revenue
uncertainty the following year. Our results demonstrate that the most effective
option for individuals to decrease revenue variability is to purchase permits
to participate in additional or more diverse fisheries. However, this option is
costly and often limited by regulations such as limited entry and catch share
programs. With increasing environmental variability in a changing climate, it
will be particularly important that individuals who rely on natural resources
for their livelihood have effective strategies to reduce income variability.

@@

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
countries have passed --- fishers must own a license to participate in
virtually all fisheries [@hilborn2003a]. While such regulations have helped
reduce overfishing, they limit the available strategies to reduce income
variability. For example, many permits limit fishing to a single species,
diversifying by purchasing additional fishing permits can be expensive (e.g.
~\$192K USD for a king and tanner crab permit in Southeast Alaska in 2016;
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some fisheries limit
the entry of new participants through individual fishing quotas (IFQs)
[@grimm2012].

Using a unique dataset collected over 30 years on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fisheries catch
species diversity on individual revenue and revenue variability that builds on
results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability.
We develop a variance-function regression model [@davidian1987; @hall1989], to
simultaneously models changes in the mean and variance of individual revenue as
a function of covariates. We simultaneously estimate the effect of species
diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
on individual revenue and revenue variability, and estimate effects both within
and across fishing strategies while controlling for fishing effort and latent
time effects (e.g. changing environmental conditions). To our knowledge, this is
the first application of this type of model in any natural resource field.

@@

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold single permits, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders --- compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
diversity of this revenue declined somewhat from 1980--2000 but has since
remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
driven by an increase in salmon specialists in the 1980s and 1990s --- the
trend disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).

@@

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut permit plus
one additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense (e.g. ~\$192K USD for a king and
tanner crab permit in Southeast Alaska in 2016
<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies with a diversity of assets provides substantial
benefits over strategies with lower diversity [REFS]. However, in practice, the
ability of individuals to effectively diversify is limited by regulation and
market forces. Here we document the effect of individual-level changes in
species diversity on revenue and revenue variability across three scales. We
find that (1) individuals with fishing permit sets that allow high levels of
species diversity tend to have lower revenue variability. (2) Adding a fisheries
permit usually increases revenue and decreases revenue variability. However, (3)
within the constraints of a permit and accounting for effort, specializing tends
to increase revenue while both specializing and generalizing tend to correlate
with greater revenue variability with considerable differences amongst
fisheries. Thus, the most effective mechanism by which species diversity can
lower revenue variability for individual fishers --- switching or adding fishing
permits --- is costly, limited by regulations, and consequently unavailable to
many individuals.

@@

We also found that altering species diversity tended to be associated with
increased revenue variability. There are number of possible reasons for this
relationship. For instance, changes to fishing patterns will sometimes be
associated with a learning or adaptation phase as individuals become accustomed
to new locations, new gear, or new species [REF]. Additionally, changes in
diversity might reflect years in which targeted species were less available and
fishers were forced to target other species or landed more bycatch adding
stochasticity to their revenue. Most importantly, this increased variability
reflects an inability for our large-scale model to predict an individual's
revenue in a subsequent year. There are many reasons, often unique to an
individual permit holder's behaviour [@hilborn1985] and therefore not included
in our large-scale model, that might make that specific individual's revenue
less predictable if species diversity changes.

>>>>

Individuals who rely on the extraction of natural resources for their livelihood
--- fishers, farmers, miners, and loggers, among others --- face high levels of
income variability driven by a mix of environmental, biological, and economic
variability. A key component to managing these industries is identifying how
regulatory actions and individual behaviour affect income variability and by
extension the economic stability and sustainable use of natural resources. For
commercial fisheries, recent work has identified that communities and vessels
that fish a greater diversity of species have less revenue variability than
those that fish a lower diversity. However, it is unclear if these benefits
extend to the level of the individual and how year-to-year changes in
diversification affect revenue and revenue variability. Here we identify three
axes by which fishers in Alaska, one of the world's most productive fishing
regions, can diversify fishing activities. We show that fishing permits with
higher species diversity reduces individual revenue variability and fishing an
additional permit is associated with higher revenue and lower variability.
However, increasing species diversity within the constraints of existing permits
has a fishery-dependent effect on revenue and is usually (`r g2_dens`%
probability) associated with increased revenue uncertainty the following year.
Our results demonstrate that the most effective option for individuals to
decrease revenue variability is to purchase permits to participate in additional
or more diverse fisheries. However, this option is costly and often limited by
regulations such as limited entry and catch share programs. With increasing
environmental variability in a changing climate, it will be particularly
important that individuals who rely on natural resources for their livelihood
have effective strategies to reduce income variability.

@@

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
countries have passed --- fishers must own a license to participate in virtually
all fisheries [@hilborn2003a]. While such regulations have helped reduce
overfishing, they limit the available strategies to reduce income variability.
For example, many permits limit fishing to a single species, diversifying by
purchasing additional fishing permits can be expensive[^1], and some fisheries
limit the entry of new participants through individual fishing quotas (IFQs)
[@grimm2012].

[^1]: For example, it cost ~\$192K USD for a king and tanner crab permit in
Southeast Alaska in 2016: <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>.

Using a unique dataset collected over 30 years on individual fishers, rather
than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fisheries catch
species diversity on individual revenue and revenue variability that builds on
results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability. We develop a variance-function regression model
[@davidian1987; @hall1989], to simultaneously model changes in the mean and
variance of individual revenue as a function of covariates. We simultaneously
estimate the effect of species diversity (the inverse of Simpson's diversity
index [@simpson1949; @jost2006]) on individual revenue and revenue variability,
and estimate effects both within and across fishing strategies while controlling
for fishing effort and latent time effects (e.g. changing environmental
conditions). To our knowledge, this is the first application of this type of
model in any natural resource field.

@@

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold a single permit, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders --- compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple-permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
diversity of this revenue declined somewhat from 1980--2000 but has since
remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely driven
by an increase in salmon specialists in the 1980s and 1990s --- the trend
disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).

@@

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\ref{fig:spoke-all}). For example, individuals fishing a halibut permit plus
one additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense.

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies that exploit a diversity of assets provides substantial
benefits over strategies with lower diversity [REFS]. However, in practice, the
ability of individuals to effectively diversify is limited by regulation and
market forces. Here we document the effect of individual-level changes in
species diversity on revenue and revenue variability across three scales. We
find that (1) individuals with fishing permit sets that allow high levels of
species diversity tend to have lower revenue variability. (2) Adding a fisheries
permit usually increases revenue and decreases revenue variability. However, (3)
within the constraints of a permit and accounting for effort, specializing tends
to increase revenue while both specializing and generalizing tend to correlate
with greater revenue variability with considerable differences amongst
fisheries. Thus, the most effective mechanism by which species diversity can
lower revenue variability for individual fishers --- switching or adding fishing
permits --- is costly, limited by regulations, and consequently unavailable to
many individuals.

@@

We also found that altering species diversity tended to be associated with
increased revenue variability. There are number of possible reasons for this
relationship. For instance, changes to fishing patterns will sometimes be
associated with a learning or adaptation phase as individuals become accustomed
to new locations, new gear, or new species [REF]. Additionally, changes in
diversity might reflect years in which targeted species were less available and
fishers were forced to target other species or landed more bycatch, adding
stochasticity to their revenue. Most importantly, this increased variability
reflects an inability for our large-scale model to predict an individual's
revenue in a subsequent year. There are many reasons, often unique to an
individual permit holder's behaviour [@hilborn1985] and therefore not included
in our large-scale model, that might make that specific individual's revenue
less predictable if species diversity changes.
